---
title: "WEEK 2 PROJECT"
author: "M.KASSEM"
date: "5/19/2019"
output: html_document
---

```{r setoptions, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide", cache = TRUE, message = FALSE, warning = FALSE)
```

## Loading libraries and Data 
```{r loading, message= FALSE}
library(ggplot2)
library(lubridate)
library(dplyr)
df <- read.csv("/Users/ayman/Downloads/R/activity.csv", header = TRUE, stringsAsFactors = FALSE)
```

## Cleaning data 
```{r cleaning}
df$date <- as.Date(as.character(df$date), "%Y-%m-%d")
```

## Data overview 
```{r overview, echo = TRUE, results= TRUE}
glimpse(df)
head(df)
```

## Total number of steps taken per day histogram 
``` {r Q2, echo = TRUE, results = TRUE}
hist(df$steps, main = "Histogram including 0 and NA values")
```

## Mean and Median steps for each day 
```{r Q3, echo= TRUE, results= TRUE}
df2 <- df %>% group_by(date) %>% summarize(mean_day = mean(steps, na.rm = TRUE), median_day = median(steps[!steps == 0], na.rm = TRUE))
head(df2)
df2n <- na.omit(df2) # removing NaN 
head(df2n, 10)
```

##  Time series plot of the average number of steps taken
```{r Q4, echo = TRUE, results= TRUE}
plot(df2$date, df2$mean_day, data = df2, pch = 20, main = "AVG steps per day time series")
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r Q5, echo = TRUE, results= TRUE}
df3 <- df %>% group_by(interval) %>% summarize(avg_steps_interval = mean(steps, na.rm = TRUE)) %>% arrange(desc(avg_steps_interval))
df3[1,]
``` 

## Code to describe and show a strategy for imputing missing data
*Missing data is 13% of the total data Therfore I am going to omit them*
```{r Q6, echo= TRUE, results= TRUE}
mean(is.na(df$steps))
```

## Histogram of the total number of steps taken each day after missing values are imputed
```{r Q7, echo= TRUE, results= TRUE}
df1 <- df %>% group_by(date) %>% summarise(steps_day = sum(steps, na.rm = TRUE))
hist(df1$steps_day, main = "Steps per day frequency")
```


## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r Q8, echo= TRUE, results= TRUE}
df4 <- df %>% group_by(week_day = wday(date, label = TRUE)) %>% summarize(mean_day = mean(steps, na.rm = TRUE))
df5 <- df4 %>% mutate(wdaylbl = case_when(week_day %in% c("Mon", "Tue", "Wed", "Thu", "Fri") ~ "Weekday",
                                          week_day %in% c("Sat", "Sun") ~ "Weekend")) %>%
        group_by(wdaylbl) %>% summarize(mean(mean_day))


as.data.frame(df5)

barplot(df5$`mean(mean_day)`, names.arg = c("Weekdays", "weekends"), ylab = "AVG steps", main = " AVG steps / 5mins intervals ")
```

